---
title: "Simple R Functions "
author: ""
date: "January 26, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}

## simple example
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b

```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2){
  
  n = length(xVec2)
  
  return(xVec2^(1:n)/(1:n))
}



c <- tmpFn2(a)
c


```

######  
  
  
  
(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$

```{R}
tmpFn3 <- function(x,n){
  return(1+sum(x^(1:n)/(1:n)))
}
```

\newpage
### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$

Try out your function.  \texttt{tmpFn}(c(1:5,6:1))

```{r}
tmpFn <- function(xVec){
  r <- c()
  r <- c(r,(xVec[1:(length(xVec)-2)]+xVec[2:(length(xVec)-1)]+xVec[3:length(xVec)])/3)
return(r)
}
tmpFn(c(1:5,6:1))
```

### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.

```{R}
tmpFn <- function(x){
  if(x<0){
    return(x^2+2*x+3)
  }
  if(x>=0 & x<2){
    return(x+3)
  }
  if(x>=2){
    return(x^2+4*x-7)
  }
}

curve(tmpFn, from=-3, to=3, xlab="x", ylab="y")
```

### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$

```{R}
doubleOdds <- function(matrix){
  x <- matrix
  x[x%%2!=0] <- x[x%%2!=0]*2
  return(x)
}
doubleOdds(matrix(c(1,1,3,5,2,6,-2,-1,-3), nrow = 3, byrow =TRUE))
```

### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$

```{R}
diagonalMatrix <- function(n,k){
  x <- matrix(rep(c(0), time=n^2), nrow = n, byrow = TRUE)
  x[abs(col(x)-row(x))==1] <- 1
  for(i in 1:n){
    for(j in 1:n){
      if(i==j){
        x[i,j] <- k
      }
    }
  }
  return(x)
}
```

### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.

```{R}
quadrant <- function(alpha){
  x <- alpha%%360
  if(x>=0 & x<90){
    return(1)
  }
  if(x>=90 & x<180){
    return(2)
  }
  if(x>=180 & x<270){
    return(3)
  }
  if(x>=270 & x<360){
    return(4)
  }
}
```

### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1`963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.

```{R}
int <- function(x){
  return(x%/%1)
}
weekday <- function(day, month, year){
  m <- month
  k <- day
  y <- year
  c <- year%%100
  return((int(2.6*m-0.2)+k+y+int(y/4)+int(c/4)-2*c)%%7)
}
```

(b) Does your function work if the input parameters \text{day, month, and year} are vectors with the same length and valid entries?

```{R}
weekday(c(1,2,3),c(12,12,12),c(2017,2017,2017))
```

Yes, the function works if the input parameters have the same length and have valid entries. 





